\textbf{Discussion On the Out-of-Sync TDC Module from the HMS Chambers}

### 2018-08-30 Spring_2018_Run/OutOfSync_TDC/tdc_sync_report.tex
final modifications to the report

\section{Observed Correlations on the Out-of-Sync TDC Module}
\noindent Figure \ref{fig:shift_dtimes} shows the Drift Chamber raw TDC times vs. Wire Number for all 12 planes. In four of the planes,
a $\sim$ 100 ns drift for cetrtain groups of wires can clearly be observed. This was the first indication, but NOT with absolute certainty,
that the TDC module to which these four groups were connected to was problematic.
\begin{figure}[h!]
  \centering
  \includegraphics[width=6.0in, height=3.0in]{shifted_drifttimes.png}
  \caption{The Raw TDC vs. Wire Number Correlations for all 12 planes. An offset of $\sim$ 1000 Channels ($\sim$ 100 ns) is observed
    in wire groups of four distinct planes.}
  \label{fig:shift_dtimes}
\end{figure}
\newpage
In Figure \ref{fig:wire_residual}, the redisuals for the tracks are plotted as a function of wire number. It can be clearly observed that the same four groups mentioned in
Figure \ref{fig:shift_dtimes} have smeared out residuals. At this point, I started to consider this issue seriously, as it was NOT simply drifted times which could be fixed in
software, but now the tarcking had been affected.
\begin{figure}[h!]
  \centering
  \includegraphics[width=7.0in, height=4.0in]{bad_wireresidual.png}
  \caption{Correlation of Residuals vs. Wire Number for all 12 planes. Four wire groups stand out from the rest, with a noticeable smearing.
  The adjacent planes are affected by this smearing as well, even though they are NOT part of the problematic TDC.}
  \label{fig:wire_residual}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[width=5.1in, height=3.1in]{hDC_RefTime_correlation.png}
  \caption{Correlation between the HMS DC reference time and the raw TDC times. The DC 1, V'-Plane raw TDC vs. Wire Number is shown. Adjacent wirgroups which
  are NOT in TDC Slot 2 are compared to the TDC Slot 2 wire group before and after applying the reference time correction to the raw TDC time.}
  \label{fig:hDC_correlation}
\end{figure}
\newpage
Figure \ref{fig:hDC_correlation} is determinant in this study, as it shows the correlation of the reference time with the raw TDC time before and after applying the reference time correction.
As expected from a normal TDC, when tdc times are reference time subtracted, the correlation dissappears. However, in TDC Slot 2, the reference time and raw tdc time appear to be
correlated, which is an indication that its copy of the reference time is Out-of-Sync with the rest of the modules.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=5.0in, height=1.0in]{loadmap.png}
  \caption{Snippet of the hms replay script that shows how to load the correct map for Spring 2018 run period.}
  \label{fig:map}
\end{figure}
\newpage
\section{How to Minimize the impact the Out-of-Sync TDC had on Residuals }
\noindent The group of wires that were affected by the TDC synchronization issue cannot be restored to normal, so their resduals will remain significantly smeared out. Fortunately,
two of the groups are located at the edge of each chamber, so they have minimal impact. The other two groups, even though they are near the center of the chamber, they are
in different chambers, so tracking in NOT severely affected. The group of wires in adjacent planes are affected (See Figure \ref{fig:wire_residual}) by this smearing, since
they depend heavily on their adjacent partners (U,V planes) for the (x,y) coordinate determination from tracking. To minimize this effect, the Hall C source code \textit{hcana}
was modified to read in the sigma parameters on a per-wire basis, if a certain flag is set. Originally, the sigma parameters were read on a per-plane basis, and were assigned a
value of 0.02 by default. If the sigma-per-wire flag is set, then the sigma parameter values are read on a wire-by-wire basis, and the wire groups associated with the
Out-of-Sync TDC are assigned a sigma of 0.06, while the rest are assigne a value of 0.02 so that during the track-fitting algorithm, if a track passes though one of this groups,
they will have a smaller contribution (or weight) to the fit. The results are shown in Figures \ref{fig:wire_residuals_flag}.
\begin{figure}[h!]
\centering
\begin{subfigure}{.5\textwidth}
  \hspace*{-5cm}                                                           
  \centering
  \includegraphics[width=2.1\linewidth]{2d_residuals_flagOFF.png}
  \subcaption{Residuals vs. Wire Number for all Chamber planes, with sigma-per-wire flag \textit{OFF}.}
  \label{fig:2d_res_flagOff}
\end{subfigure}%

>>>>

\section{Observed Correlations on the Out-of-Sync TDC Module}
\noindent Figure \ref{fig:shift_dtimes} shows the Drift Chamber raw TDC times vs. Wire Number for all 12 planes. In four of the planes,
a $\sim$ 100 ns drift for cetrtain groups of wires can clearly be observed. This was the first indication, but NOT with absolute certainty,
that the TDC module to which these four groups were connected to was problematic.
\begin{figure}[h!]
  \centering
  \includegraphics[width=6.0in, height=3.0in]{shifted_drifttimes.png}
  \caption{The Raw TDC vs. Wire Number Correlations for all 12 planes. An offset of $\sim$ 1000 Channels ($\sim$ 100 ns) is observed
    in wire groups of four distinct planes.}
  \label{fig:shift_dtimes}
\end{figure}
\newpage
In Figure \ref{fig:wire_residual}, the redisuals for the tracks are plotted as a function of wire number. It can be clearly observed that the same four groups mentioned in
Figure \ref{fig:shift_dtimes} have smeared out residuals. At this point, I started to consider this issue seriously, as it was NOT simply drifted times which could be fixed in
software, but now the tarcking had been affected.
\begin{figure}[h!]
  \centering
  \includegraphics[width=7.0in, height=4.0in]{bad_wireresidual.png}
  \caption{Correlation of Residuals vs. Wire Number for all 12 planes. Four wire groups stand out from the rest, with a noticeable smearing.
  The adjacent planes are affected by this smearing as well, even though they are NOT part of the problematic TDC.}
  \label{fig:wire_residual}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[width=5.1in, height=3.1in]{hDC_RefTime_correlation.png}
  \caption{Correlation between the HMS DC reference time and the raw TDC times. The DC 1, V'-Plane raw TDC vs. Wire Number is shown. Adjacent wirgroups which
    are NOT in TDC Slot 2 are compared to the TDC Slot 2 wire group before and after applying the reference time correction to the raw TDC time. The vertical axis
  is the HMS DC reference time corrected from itself (ns units). The horizontal axis is the raw TDC time in units of Channel ($\sim$0.1 ns/Channel).}
  \label{fig:hDC_correlation}
\end{figure}
\newpage
Figure \ref{fig:hDC_correlation} is determinant in this study, as it shows the correlation of the reference time with the raw TDC time before and after applying the reference time correction.
As expected from a normal TDC, when tdc times are reference time subtracted, the correlation dissappears. However, in TDC Slot 2, the reference time and raw tdc time
are uncorrelated, which is an indication that it is Out-of-Sync with the rest of the modules. When the raw TDC time for Slot 2 is corrected with the reference
time from another Slot, the corrected TDC time now has a correlation with the reference time which is indicative of a synchronization issue.
The reference time spans over a 25 ns range, even though it is reference time corrected from itself, which is due to the intrinsic 25 ns jitter coming from the TDC internal 40 MHz clock. The
raw TDC time spans over a $\sim$ 100 ns range, and at the smallest range, which corresponds to drift times of particles nearest to the sense wires, one can clearly see a
correlation with the reference time.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=5.0in, height=1.0in]{loadmap.png}
  \caption{Snippet of the hms replay script that shows how to load the correct map for Spring 2018 run period.}
  \label{fig:map}
\end{figure}
\newpage
\section{How to Minimize the impact the Out-of-Sync TDC had on Residuals }
\noindent The group of wires that were affected by the TDC synchronization issue cannot be restored to normal, so their residuals will remain significantly smeared out. Fortunately,
two of the groups are located at the edge of each chamber, so they have minimal impact. The other two groups, even though they are near the center of the chamber, they are
in different chambers, so tracking in NOT severely affected. The group of wires in adjacent planes are affected (See Figure \ref{fig:wire_residual}) by this smearing, since
they depend heavily on their adjacent partners (U,V planes) for the (x,y) coordinate determination from tracking. To minimize this effect, the Hall C source code \textit{hcana}
was modified to read in the sigma parameters on a per-wire basis, if a certain flag is set. Originally, the sigma parameters were read on a per-plane basis, and were assigned a
value of 0.02 by default. If the sigma-per-wire flag is set, then the sigma parameter values are read on a wire-by-wire basis, and the wire groups associated with the
Out-of-Sync TDC are assigned a sigma of 0.06, while the rest are assigne a value of 0.02 so that during the track-fitting algorithm, if a track passes though one of this groups,
they will have a smaller contribution (or weight) to the fit. The results are shown in Figures \ref{fig:wire_residuals_flag}.
\begin{figure}[h!]
\centering
\begin{subfigure}{.5\textwidth}
  \hspace*{-5cm}                                                           
  \centering
  \includegraphics[width=2.1\linewidth]{2d_residuals_flagOFF.png}
  \subcaption{Residuals vs. Wire Number for all Chamber planes, with sigma-per-wire flag \textit{OFF}.}
  \label{fig:2d_res_flagOff}
\end{subfigure}%
